Predicting the Popularity of Children’s Books Considering Reader’s Rating, Physical Attributes and Trend

### 2024-12-02 paper/paper.qmd
fixed references and rendered paper

Statistical programming language R was utilized, leveraging various R packages for efficient data management, cleaning, analysis, and visualization (@R). The tidyverse library (@tidyverse), encompassing ggplot2 and dplyr, was integral to data manipulation, aggregation, and creating visualizations tailored to the study. The janitor package (@janitor) was employed for cleaning and preparing the dataset, ensuring that variables were well-structured and analysis-ready. Additionally, broom (@broom) facilitated the conversion of statistical results into tidy formats for clear presentation. For seamless data storage and access, the here (@Here) and arrow (@arrow) libraries were used, enabling efficient file organization and handling large datasets in memory-efficient formats. Statistical modeling relied on rstanarm (@rstanarm), which allowed for Bayesian regression modeling to address the probabilistic nature of the data. Dynamic reporting was supported by the knitr package (@knitr), ensuring reproducibility and integration of code, results, and narratives into a cohesive document. The testthat (@Test) library was utilized to ensure the reliability and robustness of the code, providing a structured framework for unit testing and debugging. Furthermore, psych (@psych) was employed for descriptive statistics and exploratory data analysis, particularly for summarizing and understanding relationships between variables in the dataset.

@@

In this analysis, a linear regression model using ordinary least squares was initially considered for predicting the outcome variable. However, exploratory data analysis revealed that many variables, including the outcome, exhibit non-normal distributions typical of the exponential family. Consequently, a generalized linear model (GLM) with a log link was selected as a more suitable alternative. To compare the models, Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) were calculated for both the GLM and a linear model (LM) with a Gaussian assumption and identity link. The LM achieved slightly better error metrics, with an MSE of 1,396,426,590 and an RMSE of 37,368.79, compared to the GLM’s MSE of 1,491,689,002 and RMSE of 38,622.39.

@@

## Overview of Research and Results

This study investigates the factors influencing the popularity of children’s books by employing a generalized linear model (GLM) to predict rating counts. The dataset used includes comprehensive information about children’s books, such as their physical attributes, publication details, and user ratings, covering unique ISBNs published between 1905 and 2020. The analysis aims to understand whether attributes beyond user ratings—such as the number of pages, cover type, and republish length—contribute significantly to the popularity of children’s books. The results demonstrate that average user ratings, the number of pages, the year of publication, and the republish length are all significant predictors of popularity. Notably, average user ratings and republish length exhibit strong positive effects, suggesting that user satisfaction and renewed exposure through republication significantly drive reader engagement. While cover type shows significant effects for certain categories (e.g., "Board Books"), its overall impact is less pronounced compared to other predictors.

## Implication and Significance

Traditionally, the primary driver of a book's popularity is thought to be audience reception and average ratings (@loh2019d). However, this research challenges that conventional view by highlighting the influence of physical and temporal attributes. The findings suggest that the number of pages, a physical characteristic often overlooked, plays a meaningful role in predicting popularity. This insight provides publishers with an actionable metric—adjusting the length of books based on market trends and reader preferences could enhance engagement. Furthermore, the positive relationship between republish length and popularity emphasizes the potential of older books to regain market relevance through republication. This finding is particularly significant for publishers seeking to maximize revenue from their backlist by identifying and republishing titles with enduring appeal. Additionally, while cover type showed variability in its effects, physical formats like "Board Books" appear to hold distinct advantages in engaging readers, suggesting opportunities for targeted marketing of specific cover types.

## Research Limitation

Despite the insights gained, the research has notable limitations, primarily stemming from the dataset's scope and quality. The GLM model developed for this study showed modest predictive power, indicating that other significant predictors of book popularity remain unexamined. Variables such as genre and target audience demographics, which could provide critical context for understanding engagement, were absent from the dataset. Additionally, the dataset contained continuous variables with a high prevalence of outliers, which were retained to avoid data loss given the limited sample size. This decision may have affected model accuracy and generalizability. Moreover, the assumption of independence between observations may not fully capture complex market dynamics, such as interdependencies among books within a series or franchise.

## Conclusive Statement and Future Research

This study concludes that children’s book popularity is influenced not only by audience ratings but also by physical attributes such as the number of pages and temporal factors like republish length. Publishers can leverage these findings to make strategic decisions, such as optimizing book lengths and capitalizing on the market potential of older, republished books. However, the limitations of the current model underscore the need for further research. Future studies should include variables such as genre, target audience, and marketing efforts to enhance predictive accuracy and provide a more comprehensive understanding of popularity drivers. Addressing outliers and applying advanced machine learning techniques, such as neural networks or ensemble methods, could improve model robustness and predictive capability. Additionally, incorporating cross-validation approaches would ensure that future models are better generalized to real-world data, paving the way for more effective predictions and actionable insights in the publishing industry.

>>>>

Statistical programming language R was utilized, leveraging various R packages for efficient data management, cleaning, analysis, and visualization (@R). The tidyverse library (@tidyverse), encompassing ggplot2 and dplyr, was integral to data manipulation, aggregation, and creating visualizations tailored to the study. The janitor package (@janitor) was employed for cleaning and preparing the dataset, ensuring that variables were well-structured and analysis-ready. Additionally, broom (@broom) facilitated the conversion of statistical results into tidy formats for clear presentation. For seamless data storage and access, the here (@Here) and arrow (@arrow) libraries were used, enabling efficient file organization and handling large datasets in memory-efficient formats. Dynamic reporting was supported by the knitr package (@knitr), ensuring reproducibility and integration of code, results, and narratives into a cohesive document. The testthat (@Test) library was utilized to ensure the reliability and robustness of the code, providing a structured framework for unit testing and debugging. Furthermore, psych (@psych) was employed for descriptive statistics and exploratory data analysis, particularly for summarizing and understanding relationships between variables in the dataset.

@@

In this analysis, a linear regression model using ordinary least squares was initially considered for predicting the outcome variable. However, exploratory data analysis revealed that many variables, including the outcome, exhibit non-normal distributions consistent with the exponential family. Consequently, a generalized linear model (GLM) with a log link and Poisson distribution was selected as a more suitable alternative. The dataset was split into training and testing sets using an 80/20 split to ensure robust evaluation of model performance. The models were trained on the 80% training set, and their predictive accuracy was assessed on the 20% testing set.

To compare the models, Mean Squared Error (MSE) and Root Mean Squared Error (RMSE) were calculated for the test set. The linear model (LM) with a Gaussian assumption and identity link achieved slightly better error metrics, with an MSE of 2,506,461,514 and an RMSE of 50,064.57, compared to the GLM’s MSE of 2,514,799,221 and RMSE of 50,147.77. While the LM slightly outperforms the GLM in terms of minimizing prediction errors, the GLM with its Poisson family is better suited for the count-based outcome variable, as it aligns more closely with the data's distributional characteristics and ensures valid statistical inferences.

@@

## Overview of research and results

This study delves into the factors influencing the popularity of children’s books by employing a generalized linear model (GLM) with a log-link function to predict the count of ratings. The dataset includes extensive metadata for children’s books, such as physical attributes (number of pages and cover type), publication details (year of publication and republish length), and user feedback (average ratings and total ratings count). The data spans unique ISBNs published between 1905 and 2020, offering a wide historical range for analysis. The objective is to assess whether attributes beyond user ratings—traditionally regarded as the most significant predictor of book popularity—also play a meaningful role in determining engagement levels. Results from the analysis reveal that several factors significantly influence popularity, with average user ratings and republish length standing out as the strongest predictors. High average ratings correspond to increased engagement, reflecting the critical role of reader satisfaction in driving popularity. Republish length, defined as the time elapsed between the original publication and its republication, also exhibits a strong positive effect, highlighting the enduring relevance of older books when reintroduced to the market. The number of pages shows a moderate positive effect, suggesting that longer books may appeal to certain segments of readers, while cover type demonstrates variability in its impact. For example, "Board Books" have a distinct positive association with higher rating counts, likely reflecting their suitability for specific demographics, such as young children or educational purposes. Overall, these findings provide a nuanced understanding of the interplay between physical, temporal, and audience-related factors in shaping the popularity of children’s books.

The results underline the importance of examining a broader range of variables when predicting book popularity. While audience reception, measured through average ratings, remains a dominant factor, the study reveals that physical and temporal attributes also play significant roles. This expanded perspective challenges conventional notions that prioritize user preferences alone and introduces new dimensions for publishers to consider. For instance, republish length's positive relationship with popularity underscores the potential for publishers to strategically republish older books to capture renewed interest, especially in niche or underserved markets. Similarly, the moderate effect of the number of pages suggests that the book's length may influence its perceived value or suitability for specific purposes, such as educational or leisure reading. These insights highlight the multifaceted nature of book popularity and provide a foundation for evidence-based decision-making in the publishing industry.

## Implication and significance

Traditionally, book publishers and marketers have focused on audience reception, primarily measured through average ratings, as the key driver of popularity and market success (@loh2019d). This research challenges this singular focus by demonstrating that physical and temporal attributes of books also significantly influence reader engagement and popularity. One of the most notable findings is the positive association between the number of pages and the popularity of children’s books. While it may seem intuitive that longer books offer more content and therefore more perceived value, this study confirms that length is a measurable and actionable attribute that can influence popularity. This insight is especially valuable for publishers designing books for different reader demographics, as it suggests that optimal page lengths can be tailored to target audiences to maximize engagement. For example, longer books may appeal to older children or those seeking more immersive reading experiences, whereas shorter books may be more suitable for younger audiences.

Another critical finding is the strong positive relationship between republish length and popularity. This suggests that older books, when republished, have the potential to regain or even surpass their original popularity. This result is particularly significant for publishers looking to maximize the value of their backlist, as it highlights the opportunity to reintroduce older titles to new generations of readers. This finding aligns with broader trends in the publishing industry, where nostalgia and the rediscovery of classic books often drive market success. Additionally, the variability in the impact of cover types reveals opportunities for targeted marketing. For instance, "Board Books" show a distinct advantage in terms of popularity, likely due to their durability and suitability for young children. Publishers can leverage this knowledge to market specific formats to the appropriate audience, enhancing both reader satisfaction and sales performance. Overall, the findings provide a more comprehensive understanding of the factors driving children’s book popularity and offer actionable insights for publishers, marketers, and authors.

# Research limitation

Despite its contributions, this study has several limitations that warrant careful consideration. First, the dataset's scope and quality impose constraints on the analysis. While the GLM model used in this study identifies significant predictors, its overall predictive power is modest, suggesting that additional variables not included in the dataset could play a critical role in determining book popularity. For instance, variables such as genre and target audience demographics were absent, yet these factors are likely to have substantial effects on engagement levels. Including these variables in future studies could significantly improve model accuracy and provide a more nuanced understanding of the factors influencing popularity. Furthermore, the dataset contains continuous variables with a high prevalence of outliers, which were retained to avoid data loss given the limited sample size. While this decision preserves the dataset's integrity, it may also affect the model's accuracy and generalizability, as extreme values can disproportionately influence regression coefficients.

Another limitation lies in the assumption of independence between observations, which may not fully capture the complex market dynamics of book publishing. For example, books within the same series or franchise are likely to have interdependent popularity metrics due to shared branding, characters, or themes. Ignoring these interdependencies may lead to an oversimplification of the factors driving engagement. Additionally, the dataset does not account for external factors such as marketing efforts, seasonal trends, or the influence of awards and critical reviews, all of which could significantly impact book popularity. Addressing these limitations in future research could provide a more comprehensive and accurate understanding of the drivers of popularity, enabling publishers to make more informed decisions.

## Conclusive statement and future research

This study concludes that children’s book popularity is influenced by a combination of audience reception, physical attributes, and temporal factors. While average ratings remain a dominant predictor, attributes such as the number of pages and republish length also play significant roles in shaping reader engagement. These findings offer valuable insights for publishers seeking to optimize their strategies. For example, adjusting book lengths to align with market trends and reader preferences could enhance engagement, while strategically republishing older titles could capitalize on their enduring appeal. However, the study's limitations highlight the need for further research to build on these insights. Future studies should incorporate additional variables, such as genre, target audience, and marketing efforts, to provide a more holistic understanding of popularity drivers. Addressing outliers through robust statistical methods and applying advanced machine learning techniques, such as neural networks or ensemble methods, could improve model robustness and predictive capability.

Moreover, incorporating cross-validation approaches would ensure that future models are better generalized to real-world data, enhancing their applicability in practical settings. Future research could also explore the interactions between variables, such as how genre and cover type jointly influence popularity, to uncover deeper insights. Expanding the dataset to include international books and broader demographic information could further enrich the analysis and make the findings more generalizable. By addressing these gaps, future studies can provide more accurate predictions and actionable recommendations, ultimately benefiting publishers, authors, and readers alike. These advancements would not only improve the predictive power of models but also deepen our understanding of the multifaceted factors driving children’s book popularity in a rapidly evolving market.
